---
title: 'Linking iterative forecasting to hypothesis testing'
output:
  html_document:
    df_print: paged
bibliography: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
diagnostics <- TRUE
library(ecoforecastR)
```

Michael C. Dietze$^1$, Kathryn I. Wheeler$^1$, Colin Averill$^2$, Jennifer M. Bhatnagar$^3$, John R. Foster$^1$, Shannon L. LaDeau $^4$, Kathleen Weathers$^4$, Zoey R. Werbin$^3$,  Katherine A. Zarada$^1$

$^1$Earth and Environment, Boston University, Boston, MA

$^2$Department of Environmental Systems Sciences, ETH Zurich, Zurich, Switzerland

$^3$Department of Biology, Boston University, Boston, MA

$^4$Cary Institute of Ecosystem Studies, Millbrook, NY

# Abstract

For ESA: look into sources/citations about out-of-sample validation, experimental design, and not examining data / "sampling stopping rules” / p-hacking


paper idea: tutorial on how to perform hypothesis testing within ecological forecasting
Could start with a null model (long-term mean or random walk), propose one or more alternatives and, with a simulated data experiment, see how long it takes to refute the null (and show exactly how that’s done). Simultaneously, can assess model performance to better understand failures and propose refined models.


# Introduction

A number of recent papers have highlighted the idea of using iterative forecasting in a hypothesis testing framework as a way of accelerating learning. The proposed cycle involves using models (which embody our current hypotheses) to make out-of-sample predictions into the future, collecting data, validating results, and then assimilating this new data into models to update parameters and/or state variables. It is argued that this cycle closely mirrors the cyclic nature of the scientific method, places particular emphasis on having specific, refutable predictions, and that the validation against yet-to-be-collected data acts as preregistration and guards against overfitting. But how do you do this in practice, and how do tasks like model selection change in an iterative setting? We explore these issues using both a simulated data experiment (true model and parameters known) and a **reanalysis of a vegetation phenology forecast**. Specifically, model selection was conducted in these case studies both using an iterative approach and using **standard post-hoc model fitting, model selection, and cross-validation**. For both we used a Bayesian state-space framework, with the simulation exploring a series of dynamic linear models of increasing complexity and the phenology modeling using nonlinear logistic models with different meteorological covariates.

Q1: Is model selection valid when applied iteratively?
Q2: Does model selection change with forecast horizon?
Q3: Can we incrementally update models rather than refitting them from scratch? 


notes:

"But if we wait long enough and test after every data point, we will eventually cross any arbitrary line of statistical significance, even if there’s no real difference at all. We can’t usually collect infinite samples, so in practice this doesn’t always happen, but poorly implemented stopping rules still increase false positive rates significantly." https://www.statisticsdonewrong.com/regression.html
J. P. Simmons, L. D. Nelson, U. Simonsohn. False-Positive Psychology: Undisclosed Flexibility in Data Collection and Analysis Allows Presenting Anything as Significant. Psychological Science, 22:1359–1366, 2011.
S. Todd, A. Whitehead, N. Stallard, J. Whitehead. Interim analyses and sequential designs in phase III studies. British Journal of Clinical Pharmacology, 51:394–399, 2001.

https://en.wikipedia.org/wiki/Multiple_comparisons_problem

What is the difference between accumulation of information and *multiple comparisons*
* Data are not random -- not a new, *independent* test (e.g. cumulative meta-analysis)
* Test is cumulative, not performing a test for each individual obs.
* Not different hypotheses, same hypothesis
* multiple *sequential* tests, not *simultaneous* tests

Repeated tests, how bad can it be: http://allendowney.blogspot.com/2011/10/repeated-tests-how-bad-can-it-be.html
* scary example, but not sure I understand how they reach the conclusions they do (p-values are uniformily distributed and follow a 0-1 bound random walk?)



http://liveatthewitchtrials.blogspot.com/2011/09/ab-testing-is-khan-doing-it-wrong.html

"Now there are ways you can tell that something is statistically significant really early in a test. "Bayesian Statistics and the Efficiency and Ethics of Clinical Trials (Berry 2004 Stat Sci)" deals with these." 


"_How not to run an A/B test_ explains why and the effects of looking before all the test is finished" 
http://www.evanmiller.org/how-not-to-run-an-ab-test.html

*repeated significance testing errors*

"significance calculation makes a critical assumption that you have probably violated without even realizing it: that the sample size was fixed in advance.  If instead of deciding ahead of time, “this experiment will collect exactly 1,000 observations,” you say, “we’ll run it until we see a significant difference,” all the reported significance levels become meaningless."

"Repeated significance testing always increases the rate of false positives"

"Instead of reporting significance of ongoing experiments, report how large of an effect can be detected given the current sample size"

"Bayesian experiment design: With Bayesian experiment design you can stop your experiment at any time and make perfectly valid inferences. Given the real-time nature of web experiments, Bayesian design seems like the way forward."


Simulations showing the problems of repeated significance testing when NULL is true
```{r, echo=FALSE}
## aside: A/B testing example
pa = 0.5
pb = 0.5
nboot = 2000
nt = 2500
ABfile = "ABtest.RData"
if(file.exists(ABfile)){
  load(ABfile)
} else {
  pval <- pdiff <- na <- nb <- matrix(NA,nt,nboot)
  for(i in 1:nboot){
    na[,i] <- cumsum(rbinom(nt,1,pa)) 
    nb[,i] <- cumsum(rbinom(nt,1,pb))
    pdiff[,i] <- (na[,i]-nb[,i])/(1:nt)
  }
  tbl <- array(NA,c(nt,2,2))
  for(i in 1:nboot){
    if(i %% 100 == 0) print(i)
    tbl[,1,1] <- na[,i]
    tbl[,1,2] <- nb[,i]
    tbl[,2,1] <- (1:nt)-na[,i]
    tbl[,2,2] <- (1:nt)-nb[,i]
    for(t in 10:nt){ ## set some minimum threshold
      pval[t,i] <- chisq.test(tbl[t,,])$p.value
    }
  }
  save(pval,pdiff,file=ABfile)
}

plot(pdiff[,1],ylim=range(pdiff),type = 'n',log='x',ylab="pa - pb",xlab="Trials",xlim = c(6,nt))
for(i in seq_len(nboot)){
  lines(pdiff[,i])
}
hist(pdiff[nt,])


plot(pval[,1],ylim=c(0,1),xlim=c(10,nt),type = 'n',log='x',xlab="Trials", ylab = "P-value")
for(i in seq_len(nboot)){
  lines(pval[,i],col=i)
}
sum(apply(pval,2,min,na.rm=TRUE) < 0.05)/nboot ## cumulative rate of false positives
sum(pval[nt,] < 0.05)/nboot ## rate of false positives at end point

```



# Methods

## The null model

The most common null model for iterative timeseries is the random walk -- tomorrow is like today and any change is purely chance (stochastic process)  [[aside, for non-timeseries the most common null is a constant mean; make sure this would be included in the set of candidate models]]

$$X_{t+1} \sim N(X_t,\sigma^2)$$

Simulated data example: True model

$$X_{t+1} \sim N \left( b_1 + b_2 X_t + b_3 X_t^2 + b_4 Z_t + b_5 Z_t X_t, \sigma^2 \right)$$

```{r, echo=FALSE, cache=TRUE}
## Code to simulate first example DLM
NT <- 100
X  <- rep(NA,NT)
X[1] <- 160  ## initial condition
b <- c(-5, ## harvest
       1.5, ## 1+r
       -0.5/200, ## -r/K
       0, ## 
       0.25) ## Z/r interaction
tau <- 2 ## process error

## covariates
#pbar = 750
#psd = 250
#precip <- rgamma(NT,pbar^2/psd^2,pbar/psd^2)
Z = rnorm(NT,sin(2*pi/12*(1:NT-1)),0.2)
#plot(Z,type='l')

## forward simulation
for(t in 1:(NT-1)){
  X[t+1] <- max(0,rnorm(1,b[1] + b[2]*X[t] + b[3]*X[t]^2 + b[4]*Z[t] + b[5]*Z[t]*X[t],tau))
}

## observation error
Nplot = 1
Y = X*0
for(t in 1:NT){
  Y[t] <- mean(rpois(Nplot,X[t]))
}
if(diagnostics) plot(Y,type='l',xlab="time")
#plot(X,Y)
#plot(Z[-NT],diff(Y))
```


In the simplest case, where we assume observations are perfect, if we want to make predictions in this system there are two uncertainties we need to contend with, the process error itself and our parameteric uncertainty about the value of the process error parameter.

If we were to start from scratch fitting this model to new data...

```{r}
## non-state space fit (no obs error)
my.model = "  
  model{
  
  #### Priors
  tau ~ dgamma(a_add,r_add)
  sigma <- 1/sqrt(tau)

  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(x[t-1],tau)
  }

}"
  
#### prep data
mydat<-list(x=Y,n=length(Y),a_add=0.1,r_add=0.1)

## fit model
mc3 <- jags.model(file=textConnection(my.model),data=mydat,n.chains=3)
mc3.out <- coda.samples(model=mc3, variable.names="tau", n.iter=5000)
plot(mc3.out)

dic3 <- dic.samples(mc3,1000)

mc3.m <- as.matrix(window(mc3.out,start = 500)) ## convert to matrix, remove burn-in
tau3 <- quantile(1/sqrt(mc3.m),c(0.025,0.5,0.975))
```


As we begin to accumulate information, the parameteric uncertainty will decline asymptotically to zero, but the process error itself is irreducible, so the only way to improve predictions is to reduce the IC uncertainty. Since the process is stochastic, after each time step we would want to use new observations to figure out where we actually ended up.


```{r,results='hide'}
#### RANDOM WALK, ITERATIVE FIT
dt    <- 4
nstep <-length(Y)/dt
a_add <- r_add <- rep(0.1,nstep+1)
mp <- xp <- tauI <- matrix(NA,nstep,3)
mq <- xq <- rep(NA,nstep) #quantile of prediction
mv <- xv <- rep(NA,nstep) #variance of prediction
dicI <- matrix(NA,nstep,2)     #Iterative updating of DIC
for(i in seq_len(nstep)){

  ## prep data, pad with NA for prediction
  mydat<-list(x=c(Y[(1:dt)+(i-1)*dt],NA),n=dt+1,a_add=a_add[i],r_add=r_add[i])

  ## fit model
  mcI <- jags.model(file=textConnection(my.model),data=mydat,n.chains=3)
  mcI.out <- coda.samples(model=mcI, variable.names=c("tau",paste0("x[",dt+1,"]")), n.iter=5000)
  
  ## calculate DIC
  DIC <- dic.samples(mcI,1000)
  dicI[i,1] <- sum(DIC$deviance)
  dicI[i,2] <- sum(DIC$penalty)
  
  ## save parameters
  mcI.m <- as.matrix(mcI.out) ## convert to matrix, remove burn-in
  tauI[i,] <- quantile(1/sqrt(mcI.m[,1]),c(0.025,0.5,0.975))
  tbar <- mean(1/sqrt(mcI.m[,1]))
    
  ## update priors
  m  = mean(mcI.m[,1])
  s2 = var(mcI.m[,1])
  a_add[i+1] <- m^2/s2
  r_add[i+1] <- m/s2
  
  if(FALSE){
  hist(mcI.m,probability = TRUE)
  xseq <- seq(min(mcI.m),max(mcI.m),length=1000)
  lines(xseq,dgamma(xseq,a_add[i+1],r_add[i+1]))
  }
  
  ## save predictions
  xp[i,] <- quantile(mcI.m[,2],c(0.025,0.5,0.975)) ## prediction quantiles
  xq[i]  <- sum(mcI.m[,2] < Y[i*dt+1])/nrow(mcI.m) ## Bayes pval, predictive error
  xv[i]  <- var(mcI.m[,2])                         ## prediction variance
  
  ## prediction @ parameter mean
  mp[i,] <- Y[i*dt]+c(-1.96,0,1.96)*tbar
  mq[i]  <- pnorm(Y[i*dt+1],Y[i*dt],tbar)
  mv[i]  <- tbar^2
}
```


```{r, results='hide'}
## RANDOM WALK, full refit each chunk
ar <- rr <- rep(0.1,nstep+1)
mpr <- xpr <- tauR <- matrix(NA,nstep,3)
mqr <- xqr <- rep(NA,nstep) #quantile of prediction
mvr <- xvr <- rep(NA,nstep) #variance of prediction
dicR <- matrix(NA,nstep,2)     #Iterative updating of DIC
for(i in seq_len(nstep)){

  ## prep data, pad with NA for prediction
  mydat<-list(x=c(Y[1:(dt*i)],NA),n=i*dt+1,a_add=ar[1],r_add=rr[1])

  ## fit model
  mcR <- jags.model(file=textConnection(my.model),data=mydat,n.chains=3)
  mcR.out <- coda.samples(model=mcR, variable.names=c("tau",paste0("x[",dt*i+1,"]")), n.iter=5000)
  
  ## calculate DIC
  DIC <- dic.samples(mcR,1000)
  dicR[i,1] <- sum(DIC$deviance)
  dicR[i,2] <- sum(DIC$penalty)
  
  ## save parameters
  mcR.m <- as.matrix(mcR.out) ## convert to matrix, remove burn-in
  tauR[i,] <- quantile(1/sqrt(mcR.m[,1]),c(0.025,0.5,0.975))
  tbar <- mean(1/sqrt(mcR.m[,1]))
    
  ## update priors
  m  = mean(mcR.m[,1])
  s2 = var(mcR.m[,1])
  ar[i+1] <- m^2/s2
  rr[i+1] <- m/s2
  
  if(FALSE){
  hist(mcR.m,probability = TRUE)
  xseq <- seq(min(mcR.m),max(mcR.m),length=1000)
  lines(xseq,dgamma(xseq,ar[i+1],rr[i+1]))
  }
  
  ## save predictions
  xpr[i,] <- quantile(mcR.m[,2],c(0.025,0.5,0.975)) ## prediction quantiles
  xqr[i]  <- sum(mcR.m[,2] < Y[i*dt+1])/nrow(mcR.m) ## Bayes pval, predictive error
  xvr[i]  <- var(mcR.m[,2])                         ## prediction variance
  
  ## prediction @ parameter mean
  mpr[i,] <- Y[i*dt]+c(-1.96,0,1.96)*tbar
  mqr[i]  <- pnorm(Y[i*dt+1],Y[i*dt],tbar)
  mvr[i]  <- tbar^2
}

```


```{r}
## plot of SD: refit
tseq <- seq(dt,by=dt,length=nstep)
plot(tseq,tauR[,2],ylim=range(tauR),type='l',ylab="Standard Deviation")
ciEnvelope(tseq,tauR[,1],tauR[,3],col=col.alpha("green",0.5))
points(length(Y),tau3[2],col="red")
lines(rep(length(Y),2),tau3[c(1,3)],col="red")

## plot of SD being iteratively constrained
plot(tseq,tauI[,2],ylim=range(tauI),type='l',ylab="Standard Deviation")
ciEnvelope(tseq,tauI[,1],tauI[,3],col=col.alpha("orange",0.5))
ciEnvelope(tseq,tauR[,1],tauR[,3],col=col.alpha("green",0.5)) # refit
lines(tseq,tauR[,2],col="green")                              # refit
points(length(Y),tau3[2],col="blue")
lines(rep(length(Y),2),tau3[c(1,3)],col="blue")
legend("topright",legend=c("refit","iterative","offline"),col=c("green","orange","blue"),lwd=3) 


## plots of posterior parameters: iterative
plot(ar,col="green",type='l',lwd=3)
lines(a_add,col="orange",lwd=3)
legend("topleft",legend=c("refit","iterative"),col=c("green","orange"),lwd=3) 

plot(rr,col="green",type='l',lwd=3)
lines(r_add,col="orange",lwd=3)
legend("topleft",legend=c("refit","iterative"),col=c("green","orange"),lwd=3) 

## predictions: iterative
tpred <- seq(dt+1,by=dt,length=nstep)
plot(tpred,xp[,2],ylim=range(xp),main="One step ahead predictions")
ciEnvelope(tpred,xp[,1],xp[,3],col=col.alpha("blue",0.5))
ciEnvelope(tpred,mp[,1],mp[,3],col=col.alpha("orange",0.2))
points(tpred,Y[tpred],col="red")

## predictions: refit
tpred <- seq(dt+1,by=dt,length=nstep)
plot(tpred,xpr[,2],ylim=range(xpr),main="One step ahead predictions")
ciEnvelope(tpred,xpr[,1],xpr[,3],col=col.alpha("blue",0.5))
ciEnvelope(tpred,mpr[,1],mpr[,3],col=col.alpha("orange",0.2))
points(tpred,Y[tpred],col="red")

## quantiles
par(mfrow=c(2,2))
hist(xq)
hist(xqr)
hist(mq)
hist(mqr)
par(mfrow=c(1,1))

## predictive error partitioning
plot(tpred,pmin(1,mv/xv),type='l',main="Process Error",
     ylab="proportion predictive variance",col="orange",lwd=3)
lines(tpred,pmin(1,mvr/xvr),col="green",lwd=3)
legend("bottomright",legend=c("refit","iterative"),col=c("green","orange"),lwd=3) 

## DIC
DICi <- cumsum(apply(dicI,1,sum))
DICr <- apply(dicR,1,sum)
DIC3 <- cumsum(apply(rbind(dic3$deviance,dic3$penalty),2,sum))
plot(DIC3,type='l',ylim=range(c(DICi,DIC3,DICr)),ylab="DIC",lwd=3,col="blue")
#lines(tseq,cumsum(dicI[,1]),col="orange")
lines(tseq-1,DICi,col="orange",lwd=3)
lines(tseq-1,DICr,col="green",lwd=3)
legend("topleft",legend=c("refit","iterative","offline"),col=c("green","orange","blue"),lwd=3) 
```

While the process error is not reducible within a model, it is reducible through model improvement. Three primary ways: observation error, internal dynamics, external drivers.

Obs error: partitioning observation error from stochasticitiy in the process itself. Explain basic state space. Sampling uncertainty. observation operators.

```{r}
## state-space random
RW <- fit_dlm(model=list(obs="Y"),data=data.frame(Y=Y))
plot(RW$params)
```



Internal: could go from $N_{t+1} = N_{t}$ to $N_{t+1} = rN_{t}$ [trend, exp growth/decline] to more complex cases that would introduce stability $N_{t+1} = b_0 + b_1N_{t}$ to $N_{t+1} = b_0 + b_1N_{t} + b_2N_t^2$ etc to more complex models -- matrix models to represent different pools, age/stage classes, food webs; nonlinear models; mechanistic models, etc

External: $N_{t+1} = b_0 + b_IN_{t} + b_1Z_1 +b_2Z_2$
covariates
Special case of b_I -> 0

## Case study 1: Simple models, Offline refit

Multiple competing models represent alternative hypotheses

Want to predict with all of them? Model averaging (unweighted -> weighted). Here we want to think about not just averaging, but selection.

The simplest approach to updating, is to refit models.

Example.

```{r, results="hide"}
#foo = ParseFixed(fixed=models[[i]],cov.data = data.frame(Y=Y,Z=Z) )

SFO.file = "SFO.RData"
if(file.exists(SFO.file)){
  load(SFO.file)
} else {
  
  ## define candidate models
  models <- list()
  models[[1]] = ""
  models[[2]] = "~ X - 1"
  models[[3]] = "~ 1 + X"
  models[[4]] = "~ 1 + X + Z" 
  models[[5]] = "~ 1 + X + X^2"
  models[[6]] = "~ 1 + X + Z + X*Z"
  models[[7]] = "~ 1 + X + X^2 + Z"
  models[[8]] = "~ 1 + X + X^2 + X*Z"
  models[[9]] = "~ 1 + X + X^2 + Z + X*Z"
  
  dt=4
  nf = 12 
  nstep <-length(Y)/dt
  SFO <- list() ## simple fit, offline
  to.run = 3:9 ## list of models to run
  for(i in to.run){
    SFO[[i]] <- iterative_fit_dlm(model=list(fixed=models[[i]],obs="Y",n.iter=10000),
                                  data=data.frame(Y=Y,Z=Z),"refit",dt=dt,nf=nf)
  }
  save(SFO,Y,Z,file=SFO.file)
}
```

```{r}
## Visualize iterative forecast
for(mv in 1:length(SFO)){
  dims = dim(SFO[[mv]]$xp)
  saveGIF({
    for(t in 1:dims[1]){
      tf = seq(tseq[t],length=dims[3],by=1)
      pch = rep(1, length(Y))
      pch[1:tseq[t]] = 18
      plot(Y,type='b',xlab="time",lty=2,pch = pch)
      ciEnvelope(tf,SFO[[mv]]$xp[t,1,],SFO[[mv]]$xp[t,3,],col=col.alpha("blue",0.5))
    }
  },movie.name = paste0("fx",mv,".gif"),interval=0.5)
}
```


```{r}
#################### Conventional Analysis of case 1 ##############
 
## who won? DIC
last <- nrow(SFO[[1]]$dic) 
DIC1 <- sapply(SFO,function(x){x$dic[last,]})
DIC1 <- rbind(DIC1,apply(DIC1,2,sum))
dDIC1 <- DIC1[3,] - min(DIC1[3,])+1
plot(dDIC1,type='b',ylim=sort(dDIC1)[c(1,ceiling(length(dDIC1)*.8))]*c(0.99,1.1),
     ylab="DIC",xlab="Model",log="y")
abline(v=8,col=2,lwd=2,lty=2)
abline(v=which.min(dDIC1),col=3,lwd=2,lty=2)
legend("bottomleft",legend=c("best","true"),col=3:2,lty=2,lwd=3)

## process error
PE <- sapply(SFO,function(x){1/sqrt(as.data.frame(
                        x$paramStats[[last]]$quantiles)["tau_add",c(1,3,5)])})
plot(unlist(PE[2,]),ylim=range(PE),log="y",type='b',lwd=2,
     col="blue",xlab="model",ylab="Std. Dev. (log scale)",main="Process Error")
ciEnvelope(seq_len(ncol(PE)),PE[3,],PE[1,],col=col.alpha("blue",0.5))
abline(v=8,col=2,lwd=2,lty=2)
abline(v=which.min(unlist(PE[2,])),col=3,lwd=2,lty=2)
legend("topright",legend=c("best","true"),col=3:2,lty=2,lwd=3)

## did they get the parameters right?
extractParamCI <- function(x,p){ ## x = results, p = parameter name
  foo <- as.data.frame(t(x$paramStats[[last]]$quantiles))[c(1,3,5),p]
  if(is.null(foo)) foo = rep(NA,3)
  return(foo)
}
params <- c("tau_add","betaIntercept","betaX","betaX2","betaZ","betaX_Z")
param.true <- c(1/sqrt(tau),b)   ## convert tau to prec
param.true[3] = param.true[3]-1  ## remove RW
param.log = rep("",length(params)); param.log[1] = "y"
for(i in seq_along(params)){
  pCI <- sapply(SFO,extractParamCI,p=params[i])
  if(sum(!is.na(pCI))>0){
    yl  <- range(pCI,na.rm = TRUE)
    plot(pCI[2,],ylim=yl,type='b',lwd=2,log=param.log[i],
         col="blue",xlab="model",ylab="parameter value",main=params[i])
    polygon(c(7.6,7.6,8.4,8.4),yl[c(1,2,2,1)],col=col.alpha("green",0.5))
    for(j in 1:ncol(pCI)){
      lines(rep(j,2),pCI[c(1,3),j],col="blue",lwd=2)
    }
    abline(h=param.true[i],col=2,lwd=3,lty=2)
    abline(h=0,lty=3)
  }
}

######  **** STEP AHEAD PLOTS ARE CURRENTLY WITH THE REFIT, NOT THE FINAL MODEL *****
## one-step-ahead predictive error: plots  
tpred <- seq(dt+1,by=dt,length=nstep)
for(i in seq_along(SFO)){
  xp <- SFO[[i]]$xp[,,1]
  plot(tpred,xp[,2],ylim=range(Y),main=paste("One step ahead predictions: Model",i))
  ciEnvelope(tpred,xp[,1],xp[,3],col=col.alpha("blue",0.5))
  points(tpred,Y[tpred],col="red")
}

## one-step-ahead predictive error: RMSE
RMSE1.1p <- sapply(SFO,function(x){
  sqrt(mean((x$xp[,4,1]-Y[tpred+1])^2,na.rm = TRUE))})
plot(RMSE1.1p,ylim=range(RMSE1.1p),type='p',lwd=2,log="y",
     col="blue",xlab="model",ylab="Std. Dev. (log scale)",main="RMSE: one step ahead")
abline(v=8,col=2,lwd=2,lty=2)
abline(v=which.min(RMSE1.1p),col=3,lwd=2,lty=2)
abline(h=sd(Y),lty=2,col="grey")  ## chance
legend("topleft",legend=c("best","true","chance"),col=c(3:2,"grey"),lty=2,lwd=3)

############### LESS Conventional Analyses of case 1 ##############

## CURRENTLY WITH THE REFIT, NOT THE FINAL MODEL ##
## will differentially penalize more complex models
## need to fit final model and then add code to make n-step ahead predictions

## posterior quantiles
for(i in seq_along(SFO)){
  xq <- SFO[[i]]$xp[,6,1]
  hist(xq,main=paste("Model",i),xlim=c(0,1))
}

## n-step ahead predictive error (do you select a different model?)
RMSE1.np <- matrix(NA,nf,length(SFO))
for(i in seq_len(nf)){
  RMSE1.np[i,] <- sapply(SFO,function(x){
                    sqrt(mean((x$xp[,4,i]-Y[tpred+i])^2,na.rm = TRUE))})
}
plot(RMSE1.np[,1],ylim=c(min(RMSE1.np),2*sd(Y)),lwd=3,
     log='y',type='l',xlab="lead time",
     ylab="Std. Dev. (log scale)",main="RMSE: n-step ahead")
for(i in 2:length(SFO)){
  lines(RMSE1.np[,i],col=i,lwd=3)
}
abline(h=sd(Y),lty=2,col="grey")
#abline(v=8,col=2,lwd=2,lty=2)
#abline(v=which.min(RMSE1.1p),col=3,lwd=2,lty=2)
sel = which(RMSE1.1p < sd(Y))
legend("topleft",legend=paste("M",sel),col=sel,lwd=3)


## risk of false positives

```

Conclusions:
* predictive error (RMSE) is probably the best overall metric
    * should also be reflected in Deviance?
* "best" model changes with lead time

```{r}
#################### Time-evolving Analysis of case 1 ##############

## Deviance vs time
DIC1.t <- matrix(NA,length(SFO),length(tpred))
for(i in seq_along(SFO)){
  DIC1.t[i,] <- apply(SFO[[i]]$dic,1,sum)/(tpred-1)
}
DIC1.t = DIC1.t - min(DIC1.t) + 1
plot(DIC1.t[1,],type='n',ylim=c(min(DIC1.t),quantile(DIC1.t,0.8)),
     ylab="DIC/n",xlab="Model",log="y")
for(i in seq_along(SFO)){
  lines(DIC1.t[i,],col=i)
}
sel <- seq_along(SFO)
legend("topright",legend=paste("M",sel),col=sel,lwd=2,horiz=TRUE,cex=0.7)

## Model weight (Deviance) vs time

## RMSE at different lags vs time
RMSE1.te <- array(NA,c(length(SFO),nf,length(tpred))) ## te = time evolving
for(m in seq_along(SFO)){
  for(i in seq_len(nf)){
    for(t in 1:length(tpred)){
      RMSE1.te[m,i,t] <- sqrt(mean((SFO[[m]]$xp[1:t,4,i]-Y[tpred[1:t]+i])^2,na.rm = TRUE))
    }
  }
}
RMSE1.te[is.nan(RMSE1.te)] <- NA
n = seq_along(tpred)
qlo = 1/sqrt(qgamma(0.975,n/2,n/2))
qhi = 1/sqrt(qgamma(0.025,n/2,n/2))
for(j in 1:nf){
  plot(tpred,RMSE1.te[1,j,],ylim=c(min(RMSE1.te,na.rm = TRUE),2*sd(Y)),lwd=3,
       log='y',type='l',xlab="time",
       ylab="Std. Dev. (log scale)",main=paste0("RMSE evolution: ",j,"-step ahead"))
  for(i in 1:length(SFO)){
     lines(tpred,RMSE1.te[i,j,],col=i,lwd=5)
     ciEnvelope(tpred,RMSE1.te[i,j,]*qlo,RMSE1.te[i,1,]*qhi,col=col.alpha(i,0.25))
  }
  abline(h=sd(Y),lty=2,col="grey")
  sel = which(RMSE1.1p < sd(Y))
  legend("topleft",legend=paste("M",sel),col=sel,lwd=3,horiz = TRUE)
}



## predictive variance vs time, multiple lags
for(i in seq_len(nf)){
  plot(tpred,tpred,type='n',ylim=c(5,2*sd(Y)),lwd=3,
       log='y',xlab="time", ylab="Std. Dev. (log scale)",
       main=paste0("Predictive Error evolution: ",i,"-step ahead"))
  for(m in seq_along(SFO)){
    lines(tpred,sqrt(SFO[[m]]$xp[,5,i]),col=m,lwd=3)
  }
  abline(h=sd(Y),lty=2,col="grey")
  sel = seq_along(SFO)
  legend("topleft",legend=paste("M",sel),col=sel,lwd=3,cex=0.7,horiz = TRUE)
}

## predictive loss
for(i in seq_len(nf)){
  plot(tpred,tpred,type='n',ylim=c(5,4*sd(Y)),lwd=3,
       log='y',xlab="time", ylab="Std. Dev. (log scale)",
       main=paste0("Predictive Loss evolution: ",i,"-step ahead"))
  for(m in seq_along(SFO)){
    lines(tpred,sqrt(SFO[[m]]$xp[,5,i]+RMSE1.te[m,i,]^2),col=m,lwd=3)
  }
#  abline(h=sd(Y),lty=2,col="grey")
  sel = seq_along(SFO)
  legend("topleft",legend=paste("M",sel),col=sel,lwd=3,cex=0.7,horiz = TRUE)
}

## Model weight  vs time, multiple lags?
MW1.te <- array(NA,c(length(SFO),nf,length(tpred)))
for(i in seq_len(nf)){
  for(m in seq_along(SFO)){ ## calculate PL precisions
    MW1.te[m,i,] <- 1/(SFO[[m]]$xp[,5,i]+RMSE1.te[m,i,]^2)
  }
  for(t in seq_along(tpred)){ ## normalize
    MW1.te[,i,t] = MW1.te[,i,t]/sum(MW1.te[,i,t])
  }
}
for(i in seq_len(nf)){
 barplot(MW1.te[,i,],col=1:length(SFO), xlim=c(1, length(tpred) + 6.5),
         main=paste0("PL model weight evolution: ",i,"-step ahead")) 
 legend(10,0.5,legend=paste0("M",sel),fill=sel,cex=0.7,
        x=length(tpred)+4.75, y=1.0,bty='n')
}


```


## Part 2 -- What about the false positive problem?

* Compare RW to univariate DLM via numerical simulation (many simulated data sets)
  * how often do we select DLM when RW is 'true'?
  
* linear DLM to quadratic DLM (logistic)

* internal only vs covariates



## Case study 2: Simple models, Iterative update


```{r}

```


## Case study 3: Iterative update + adaptive UA: sampling

## Case study 4: Iterative update + adaptive model analysis (model proposal; branching paths)

# Results

For choosing among competing models, we employ an out-of-sample predictive loss metric, which aims to balance validation error and total predictive uncertainty (parameter, initial condition, driver, and process errors). In an iterative context, all predictive uncertainties are initially at their maximum (prior) values, and thus the simplest model (random walk) initially dominates. After 10-20 time steps, parameter and process error estimates begin to converge for simple models and we are able to begin selecting among alternative model structures and covariates. Model selection takes longer for the phenological models, where the key information comes from once-a-year-transitions, and thus choosing between hypotheses requires multiple years of data regardless of measurement frequency. In addition to simple iterative model selection, we also assessed the ability of iterative validation metrics to help diagnose potential covariates and models structural errors. Finally, posthoc data analysis consistently selected for more complex model structures, and had a higher rate of overfitting than the iterative approach. Similarly, we found that multi-model iterative approaches need to retain alternative models beyond the point when a candidate model is ‘significant’ to prevent the premature ‘chance’ convergence on incorrect models.

# Discussion

# Acknowledgements

# References



